On the discovery and origin of $r$-process stars in the Milky Way

### 2018-02-06 article/ms.tex
Final final draft
intro + abstract comments

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[On the Galactic nucleosynthesis of $r$-process material.]{On the Galactic nucleosynthesis of $r$-process material.}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Matthew T. Miles et al.]{Matthew T. Miles,$^{1}$\thanks{E-mail: mtmil3@student.monash.edu}
	Andrew R. Casey,$^{1,2}$
	Brodie J. Norfolk,$^{1}$
	Alex J. Kemp,$^{1}$\newauthor
	John C. Lattanzio,$^{1}$
	Kevin C. Schlaufman,$^{3}$
	Alexander P. Ji,$^{4}$
	Anna Y. Q. Ho,$^{5}$
	\\
	% List of institutions
	$^{1}$School of Physics and Astronomy, Monash University, Clayton Campus, Victoria 3800, Australia\\
	$^{2}$Faculty of Information Technology, Monash University, Clayton Campus, Victoria 3800, Australia\\
	$^{3}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N Charles St., Baltimore, MD 21218, USA\\
	$^{4}$Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA\\
	$^{5}$Cahill Center for Astrophysics, California Institute of Technology, MC 249-17, 1200 E California Blvd, Pasadena, CA, 91125, USA
}

@@

	\begin{abstract}
		The production of heavy elements occurs through the slow and rapid neutron capture process. Numerous sites satisfy the conditions needed to synthesise r-process elements, but the frequency and abundance of these sites is unclear. The lack of a large number of stars with strong r-process signatures limits us. Here we report the discovery of 47 r-process enhanced stars from the LAMOST survey, more than doubling the known number in our Galaxy. Additionally, we provide the first estimation, to our knowledge, of whether neutron star mergers are able to act as the sole site for r-process nucleosynthesis with regards to all known r-process enhanced stars in the Milky Way. The discovery of these stars implies multiple r-process sites, contrary to previous conclusions in the literature including all r-process material being produced in small amounts in common events such as core-collapse supernovae, or the material exclusively being produced in very large amounts in very rare events, such as neutron star mergers. 
	\end{abstract}

@@

	For over 60 years theorists have argued about the potential sites of the r-process. Rapid neutron capture occurs when the neutron density is high enough that the rate of neutron capture will occur faster than the rate at which the isotope undergoes a $\beta$-decay. The rapid capture sites are very specific, with the most favoured being core-collapse supernovae, neutron star mergers, or magnetorotationally driven supernovae. Core-collapse supernovae are relatively common throughout the Galaxy, but they only produce small amounts of r-process material. Neutron star mergers or magnetorotationally driven supernovae, however, occur much more infrequently than core-collapse supernovae, but produce far more r-process material. Events such as these are theorised to account for the case of the highly r-process enhanced ultra-faint dwarf (UFD) galaxy Reticulum II \citep{Ji2016}. Neutron star mergers may have a stronger argument to be the parent site of this level of r-process enhancement in such a small volume. However, magnetorotationally driven supernovae are also a possible solution.

	While magnetorotationally driven supernovae may not produce as much r-process material as neutron star mergers (although still more than core-collapse supernovae), they occur more frequently, and enough material is produced that it is currently impossible to distinguish between them through abundance signatures.

	Europium (Z=63), is entirely produced through the r-process. This makes it a very useful signature when detecting stars that show evidence of pollution by r-process material. The recent discovery of the UFD Reticulum II used Europium as evidence of its r-process enhancement, and found the particularly interesting result that the UFD's level of r-process enrichment was too high to have occurred through a core-collapse supernovae event.

	Historically UFD's have had some of the lowest abundance of r-process elements of any celestial body. However analysis of several of Reticulum II's stars show they are greatly enriched, 2-3 orders of magnitude higher than stars found in any other UFD galaxy \citep{Ji2016}. This degree of enrichment supports the theory that a single rare event must have occurred in the galaxies formative years. This is supported by the observed heavy element yields (Europium especially) which are found to be 1000 times higher than are achieved from core-collapse supernovae ejecta. It is highly improbable that 1000 supernovae could have contributed to Reticulum II, then the enrichment of this UFD galaxy suggests that some r-process material must be produced by a rare astrophysical event such as a neutron star-neutron star merger.

	While neutron star mergers or magnetorotationally driven supernovae are likely to produce significant amounts of r-process material. However, this is not to suggest that all r-process material stems from these sources, rather, the rarity of these events possibly indicates the rarity of stars found to be heavily enhanced in these elements. What we then have to ask is can these rare events be responsible for all r-process stars found in the Milky Way, regardless of the relative strength of their abundance? Or perhaps, assuming very little mixing, could core-collapse supernovae create enough material to pollute the stars we've found? And if neither of these prove to be the case, can we then conclude that there must be multiple sites of r-process element creation throughout the Milky Way, that create these elements to different extents?

	To answer this question it is necessary to find and catalogue more r-process stars to find some relative fraction of enhanced stars that are either very enhanced, and so must come from a rare event. Or only slightly enhanced, and so must come from a more common event that produces less material. However, r-process stars are very rare, which makes finding this fraction difficult. For this reason we must look to massive data sets, such as the LAMOST sky survey.

	In this letter we report the probable discovery of 47 of these LAMOST giants that show enhanced r-process peaks for Europium at both 4129 and 4205 Angstroms.
	In Section 2 we outline the methods used to attain our results and the analysis done. In Section 3 we discuss what our results imply about the r-process. We conclude in Section 4.

@@

	\bibliographystyle{mnras}
	%\bibliography{paper} % if your bibtex file is called example.bib
	\begin{thebibliography}{99}
		\bibitem[\protect\citeauthoryear{Sneden, C. et al.}{2008}]{Sneden2008}
		Sneden C., Cowan J.J., \& Gallino R. 2013, Annual review of Astronomy \& Astrophysics, 46, 1
		\bibitem[\protect\citeauthoryear{Ji, A. P. et al.}{2016}]{Ji2016}
		Ji A. P., Frebel A., Chiti A., \& Simon J. D. 2016, Nature, 531
		\bibitem[\protect\citeauthoryear{Ho et al.}{2017}]{AnnaHo2017}
		Ho A. Y. Q., Ness M. K., Hogg David W., Rix H-W., Liu C., Yang F., Zhang Y., Hou Y.,\& Wang Y. 2017, ApJ, 836, 1
		\bibitem[\protect\citeauthoryear{Burbidge, E. et al.}{1957}]{Burbidge1957}
		Burbidge E. M., Burbidge G. R., Fowler W. A., \& Hoyle F. 1957, Review of Modern Physics, 29, 4
		\bibitem[\protect\citeauthoryear{Goriely, S. et al.}{2011}]{Goriely2011}
		Goriely S., Bauswein A. \& Janka H-T. 2011, , 738, L32
		\bibitem[\protect\citeauthoryear{Abbott, B.P. et al.}{2016}]{LIGO2016}
		Abbot B.P. et al. 2016, Living Reviews in Relativity, 19, 1
		\bibitem[\protect\citeauthoryear{Argast, D. et al.}{2004}]{Argast2004}
		Argast D., Samland M., Thielmann F-K. \& Qian Y-Z. 2004, Astronomy and Astrophysics, 416
		\bibitem[\protect\citeauthoryear{Woosley, S.E. et al.}{1992}]{Woosley1992}
		Woosley S.E, \& Hoffman R.D. 1992, ApJ, 395, 1
		\bibitem[\protect\citeauthoryear{Burrows, A. et al.}{1995}]{Burrows1995}
		Burrows, A., Hayes, J., \& Fryxell, Bruce A. 1992, ApJ, 450, 830
		\bibitem[\protect\citeauthoryear{Li, W. et al.}{2011}]{Li2011}
		Li, W., Chornock, R., Leaman, J., Filippenko, A. V., Poznanski, D., Wang, X., Ganshalingam, M., \& Mannucci, F. 2011, Royal Astronomical Society, 412, 1473
		\bibitem[\protect\citeauthoryear{Drout, M. R. et al.}{2017}]{Drout2017}
		Drout, M. R. et al. 2017, Science
		\bibitem[\protect\citeauthoryear{Kasen, D. et al.}{2017}]{Kasen2017}
		Kasen, D., Metzger, B., Barnes, J., Quataert, E., \& Ramirez-Ruiz, E. 2017, Nature, 551, 7678
		\bibitem[\protect\citeauthoryear{Hotokezaka, K. et al.}{2013}]{Hotok2013}
		Hotokezaka, K., Kiuchi, K., Kyutoku, K., Okawa, H., Skiguchi, Y., Shibata, M., \& Taniguchi, K. 2013, APS
	\end{thebibliography}

@@

			-               	& {[}deg{]}   & {[}deg{]}	& {[}$\rm{pixel}^{-1}${]} & {[}$\rm{km\ s}^{-1}${]} & {[}K{]}             	& {[}$cm\ s^{-2}${]} & {[}dex{]}          	& {[}dex{]}              	& -                        	& {[}dex{]}        	& {[}dex{]}  	\\ \midrule

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Discovery and origin of $r$-process stars]{On the discovery and origin of $r$-process stars in the Milky Way}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Matthew T. Miles et al.]{Matthew T. Miles,$^{1}$\thanks{E-mail: mtmil3@student.monash.edu}
	Andrew R. Casey,$^{1,2}$
	Brodie J. Norfolk,$^{1}$
	Alex J. Kemp,$^{1}$\newauthor
	John C. Lattanzio,$^{1}$
	Kevin C. Schlaufman,$^{3}$
	Alexander P. Ji,$^{4}$
	Anna Y. Q. Ho$^{5}$
	\\
	% List of institutions
	$^{1}$School of Physics and Astronomy, Monash University, Clayton Campus, Victoria 3800, Australia\\
	$^{2}$Faculty of Information Technology, Monash University, Clayton Campus, Victoria 3800, Australia\\
	$^{3}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N Charles St., Baltimore, MD 21218, USA\\
	$^{4}$Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA\\
	$^{5}$Cahill Center for Astrophysics, California Institute of Technology, MC 249-17, 1200 E California Blvd, Pasadena, CA, 91125, USA
}

@@

	\begin{abstract}
		The production of heavy elements occurs through the slow and rapid neutron capture process. Numerous sites satisfy the conditions needed for the r-process to occur, but the relative frequency of these sites is unknown largely due to the lack of a large number of stars with strong r-process signatures. Here we report the discovery of 47 r-process enhanced stars discovered in the LAMOST survey, more than doubling the known number in our Galaxy. To our knowledge, we provide the first estimate on whether neutron star mergers are frequent enough to account for all known r-process stars in the Milky Way. \todo{We conclude....}. \todo{The discovery of these stars implies multiple r-process sites.}
		% contrary to previous conclusions in the literature including all r-process material being produced in small amounts in common events such as core-collapse supernovae, or the material exclusively being produced in very large amounts in very rare events, such as neutron star mergers. 
	\end{abstract}

@@

	Potential sites of the r-process have been discussed for over sixty years. Rapid neutron capture occurs when the neutron density is high enough that the rate of neutron capture will occur faster than the rate at which the isotope undergoes a $\beta$-decay. The rapid neutron capture sites are very specific, with the most favoured being core-collapse supernovae, neutron star mergers, or magnetorotationally driven supernovae. Core-collapse supernovae are relatively common throughout the Galaxy, but they only produce small amounts of r-process material. Neutron star mergers or magnetorotationally driven supernovae, however, occur much less frequently than core-collapse supernovae, but produce far more r-process material. Events such as these are theorised to account for the case of the highly r-process enhanced ultra-faint dwarf (UFD) galaxy Reticulum II \citep{Ji2016}. %Neutron star mergers may have a stronger argument to be the parent site of this level of r-process enhancement in such a small volume. However, magnetorotationally driven supernovae are also a possible solution.
Magnetorotationally driven supernovae may not produce as much r-process material as neutron star mergers (although still more than core-collapse supernovae), they are thought to occur more frequently than neutron-star mergers \todo{[citation needed]}, and the predicted abundance patterns make them observationally indistinguishable from neutron star mergers \citep{Ji2016}.

	Europium ($Z=63$), is entirely produced through the r-process. This makes it a very useful signature for identifying stars that show evidence of excess pollution by r-process material. %\todo{CLUMSY, lose the first half-sentence:The recent discovery of the UFD Reticulum II used Europium as evidence of its r-process enhancement, and found the particularly interesting result that the UFD's level of r-process enrichment was too high to have occurred through a core-collapse supernovae event.}
	Historically, UFD galaxies have demonstrated some of the lowest abundance ratios of r-process elements seen in the Milky Way or its satellite systems. However, the analysis of several of Reticulum II's stars show they are greatly enriched, 2-3 orders of magnitude higher than stars found in any other UFD galaxy \citep{Ji2016}. This degree of enrichment supports the theory that a single rare event must have occurred early on. This is supported by the observed heavy element yields which are found to be some 1000 times higher than what is achievable on average from core-collapse supernovae ejecta. It is highly improbable that 1000 supernovae could have contributed to Reticulum II, suggesting that the r-process material must be produced by a rarer astrophysical event such as a binary neutron star merger.

	While neutron star mergers or magnetorotationally driven supernovae are likely to produce significant amounts of r-process material, it does not suggest that all r-process material stems from these sources either. Rather, the rarity of these events possibly indicates the rarity of stars found to be heavily enhanced in these elements. One could ask if neutron-star mergers could be responsible for all r-process stars known in the Milky Way? On the other hand, if we assume very little gas mixing, could could core-collapse supernovae create enough material to explain a single r-process star? If neither of these prove to be true then we can confirm that there must be multiple sties to the r-process, providing a step forward in understanding the relative fraction of these mechanisms in the Milky Way.

%	 stars we've found? And if neither of these prove to be the case, can we then conclude that there must be multiple sites of r-process element creation throughout the Milky Way, that create these elements to different extents?

	To answer these questions it is necessary to identify a large sample of r-process stars.
	 However, r-process stars are very rare, which makes finding this stars difficult. For this reason we must look to massive data sets, such as the LAMOST sky survey. In this letter we report the  discovery of 47 r-process enhanced stars identified from LAMOST.
	In Section 2 we outline the methods used to identify r-process enhanced candidates, and our analysis methods. In Section 3 we discuss what our results imply about the sites of r-process. We provide concluding remarks in Section 4.

@@

	\bibliographystyle{mnras}
	%\bibliography{paper} % if your bibtex file is called example.bib
	\begin{thebibliography}{99}
		\bibitem[\protect\citeauthoryear{Abbott, B.P. et al.}{2016}]{LIGO2016}
		Abbot B.P. et al. 2016, Living Reviews in Relativity, 19, 1
		\bibitem[\protect\citeauthoryear{Argast, D. et al.}{2004}]{Argast2004}
		Argast D., Samland M., Thielmann F-K. \& Qian Y-Z. 2004, Astronomy and Astrophysics, 416
		\bibitem[\protect\citeauthoryear{Burbidge, E. et al.}{1957}]{Burbidge1957}
		Burbidge E. M., Burbidge G. R., Fowler W. A., \& Hoyle F. 1957, Review of Modern Physics, 29, 4
		\bibitem[\protect\citeauthoryear{Burrows, A. et al.}{1995}]{Burrows1995}
		Burrows, A., Hayes, J., \& Fryxell, Bruce A. 1992, ApJ, 450, 830
		\bibitem[\protect\citeauthoryear{Drout, M. R. et al.}{2017}]{Drout2017}
		Drout, M. R. et al. 2017, Science
		\bibitem[\protect\citeauthoryear{Goriely, S. et al.}{2011}]{Goriely2011}
		Goriely S., Bauswein A. \& Janka H-T. 2011, , 738, L32
		\bibitem[\protect\citeauthoryear{Ho et al.}{2017}]{AnnaHo2017}
		Ho A. Y. Q., Ness M. K., Hogg David W., Rix H-W., Liu C., Yang F., Zhang Y., Hou Y.,\& Wang Y. 2017, ApJ, 836, 1
		\bibitem[\protect\citeauthoryear{Hotokezaka, K. et al.}{2013}]{Hotok2013}
		Hotokezaka, K., Kiuchi, K., Kyutoku, K., Okawa, H., Skiguchi, Y., Shibata, M., \& Taniguchi, K. 2013, APS
		\bibitem[\protect\citeauthoryear{Ji, A. P. et al.}{2016}]{Ji2016}
		Ji A. P., Frebel A., Chiti A., \& Simon J. D. 2016, Nature, 531
		\bibitem[\protect\citeauthoryear{Kasen, D. et al.}{2017}]{Kasen2017}
		Kasen, D., Metzger, B., Barnes, J., Quataert, E., \& Ramirez-Ruiz, E. 2017, Nature, 551, 7678
		\bibitem[\protect\citeauthoryear{Li, W. et al.}{2011}]{Li2011}
		Li, W., Chornock, R., Leaman, J., Filippenko, A. V., Poznanski, D., Wang, X., Ganshalingam, M., \& Mannucci, F. 2011, Royal Astronomical Society, 412, 1473
		\bibitem[\protect\citeauthoryear{Sneden, C. et al.}{2008}]{Sneden2008}
		Sneden C., Cowan J.J., \& Gallino R. 2013, Annual review of Astronomy \& Astrophysics, 46, 1
		\bibitem[\protect\citeauthoryear{Woosley, S.E. et al.}{1992}]{Woosley1992}
		Woosley S.E, \& Hoffman R.D. 1992, ApJ, 395, 1
	\end{thebibliography}

@@

			-               	& {[}H:M:S{]}   & {[}H:M:S{]}	& {[}$\rm{pixel}^{-1}${]} & {[}$\rm{km\ s}^{-1}${]} & {[}K{]}             	& {[}$cm\ s^{-2}${]} & {[}dex{]}          	& {[}dex{]}              	& -                        	& {[}dex{]}        	& {[}dex{]}  	\\ \midrule
